Reproducible Research in R: A tutorial on how to do the same thing more than once

### 2021-09-29 manuscript.Rmd
typo

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
The comparison is made impossible if the authors of the preregistration were intentionally vague.
Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

@@

We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.

Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
In fact, we expect that most PACs will require changes after data have been collected.
Often deviations provide an opportunity to learn from the unexpected.
Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].

Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
However, after fitting the model on the actual data, both constructs exhibit high collinearity.
The authors decide that for their purpose, these variables are interchangeable and remove one.
For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
Another researcher, however, who, e.g. hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
With transparent reporting, it is left to the reader to make up their own judgement.

@@

Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier2021].
These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
We specifically recommend PAC for those exploratory statistical models.
The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.

>>>>

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

@@

We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
Each entry in the change log should contain the commit id to the respective changes.
The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.

Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
In fact, we expect that most PACs will require changes after data have been collected.
Often deviations provide an opportunity to learn from the unexpected.
Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].

Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
However, after fitting the model on the actual data, both constructs exhibit high collinearity.
The authors decide that for their purpose, these variables are interchangeable and remove one.
For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
Another researcher, however, who, e.g. hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
With transparent reporting, it is left to the reader to make up their own judgement.

@@

Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning. <!--citation? reproducibility in machine learning research? maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models? -->
These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
We specifically recommend PAC for those exploratory statistical models.
The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
